### 2024-04-15 src/tex/ms.tex
Finish text

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{Spirals launched by a planet in a shadowed disk}

% Author list
\author{Ted Johnson}
\affiliation{UNLV}

\newcommand{\mpl}{10^{-5}}

\begin{abstract}
    The presence of a planet in a protoplanetary disk is known to launch pressure and
    density spirals in the disk. A disk that is shadowed (via an optically thick inclined inner disk)
    will change the shape of these spirals as it modulates the speed of sound in the outer disk. The presence of a 
    shadow also has major effects independent of the planet. We explore these effects using 2D disk simulations with
    the MHD code Athena++.
\end{abstract}

\section{Introduction}
\label{sec:intro}

A planet embedded in a protoplanetary disk will launch a spiral similar to those seen in galaxies.
Sound waves excited by the planet travel in the rotating disk at the local sound speed, but also follow
the Keplerian rotation of the disk material. This creates a spiral shape, as waves launched interior to
the planet outpace it and waves launched outside lag behind.

Most simulations of protoplanetary disks assume that the sound speed is constant throughout the disk,
however this may not always be the case. An interior, truncated disk that is misaligned to the main disk can
cast shadows on its surface, locally lowering the temperature and sound speed. Shadows can have two effects:
\begin{enumerate}
    \item Deflecting waves launched by a planet (similar to Snell's Law).
    \item Launching its own waves because the disk now has a pressure gradient in the $\phi$ direction.
\end{enumerate}

Both effects will affect an observation of a planet in a shadowed disk. In this paper we will describe our simulation
of such a shadowed disk in Section \ref{sec:methods}, discuss the effects introduced by the shadow in Section \ref{sec:results},
and suggest future investigations in section \ref{sec:conclusion}

\section{Methods}
\label{sec:methods}
\subsection{Simulation setup}
\label{subsec:setup}
We use the magneto-hydrodynamics code Athena++\footnote{\url{https://github.com/PrincetonUniversity/athena}} 
\citep{stone2020} version 21.0 \citep{athena++developmentteam2021} to simulate a 2D gas disk. The simulation was set up in cylindrical coordinates with 149
cells in the $\ln{r}$ direction with $r \in [0.4, 2.5]$ and 512 cells in the $\phi \in [-\pi, \pi]$ direction. The inner and outer boundaries have a user-defined fixed boundary
condition and the azimuthal boundary has a periodic boundary condition to ensure material can orbit the central star indefinitly. The dimensionality of the problem
is set up so that all masses are in $M_\odot$, distances are in AU, and times are in year/$2\pi$. This means that a planet placed in a circular orbit at $r=1$
will have an orbital period of $2\pi$. We chose a planet mass of $\mpl$ in order to create a detectable effect in the presence of noise from the shadow and
to not stray too far from the linear spiral regime.

The disk is set up so that $\rho = 1$ at $r=1$ and has the radial profile $\rho \propto r^{-1}$. The gas in the disk initially moves at the local keplerian velocity.
A shadow on the disk is prescribed by setting the sound speed -- analogous to setting the temperature as 
\begin{equation}
    c_s = \frac{\partial P}{\partial \rho} \propto T
\end{equation}
We ensure that the shadow does not move with the gas in its orbit by setting the cooling timescale to $10^{-6} \ll 1$. Essentially, the temperature is reset each time
the grid is updated.

The shadow is parameterized by three numbers: \texttt{h\_reduced\_factor}, \texttt{half\_shadow}, and \texttt{half\_unshadow},
which we will write here as $h$, $a$, and $b$, respectively. The former is the dimensionless magnitude of the sound speed reduction
and the latter two are the half-width of the regions with constant sound speed (i.e. not the transition region) and have
units of radians.

The sound speed outside the shadowed region can be computed
\begin{equation}
    c_{\rm un} = c_0 r^{-1}
\end{equation}

where $c_0 = 0.1$ is the sound speed at $r=1$. The effect due to the shadow is parameterized by the equation
\begin{equation}
    c_{\rm sh} = c_{\rm un} H(\phi)
\end{equation}

where

\begin{equation*}
    \label{eq:H}
    \begin{array}{cccc}
        H = & h & \text{if} & |\sin{\phi}| \le \sin{a}  \\
            & 1 & \text{if} & |\cos{\phi}| \le \sin{b} \\
            & h + \frac{1-h}{2} \left (  1-\cos{\left ( \frac{2 \pi | \phi + m\pi |}{\pi - 2a - 2b}  \right )  } \right ) & & \text{otherwise}
    \end{array}
\end{equation*}
and
\begin{equation}
    \begin{array}{cccc}
        m = & 1 & \text{if} & \phi < -\frac{\pi}{2} \\
            & -1 & \text{if} & \phi > \frac{\pi}{2} \\
            & 0 & & \text{otherwise}
    \end{array}
\end{equation}
Figure \ref{fig:cs} shows the sound speed profiles used in this work.

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5in]{figures/sound_speed.pdf}
        \caption{Sound speed as a function of azimuthal angle. Here we show how the sound speed is prescribed in our two science cases. The cases
        share a minimum and maximum sound speed, but the transitions between these values are very different.}    
    \end{center}
    \script{plot_sound_speed.py}
    \label{fig:cs}
\end{figure}

\begin{figure*}
    \includegraphics[width=1.0\textwidth]{figures/initial_temperature.pdf}
    \caption{Initial gas temperature maps for our three setups. Note that the radial axis is scaled logarithmically. {\bf Left:} In the case where there is no shadow the temperature is azimuthally symmetric.
    {\bf Center:} In our ``narrow'' case we see a near-constant shadow covers a narrow region of the disk. This is analogous to a thin, optically thick inner disk.
    {\bf Right:} In our ``wide'' case we see the shadowing is more gradual and covers a significant portion of the disk. In this case $a,b=0$ and the third line of
    Equation \ref{eq:H} becomes a sine function for all $\phi$.
    }
    \label{fig:setup}
    \script{plot_initial.py}
\end{figure*}

We will investigate two scenarios:
\begin{enumerate}
    \item A narrow shadow, analogous to that produced by a thin inner disk with a scale height of $\frac{h}{r} = 0.05$. We chose a value of $a$ based on the assumption that the inner disk
        is optically thick up to 3 scale heights. We call this case ``narrow''.
    \item A wide shadow, like that produced by a diffuse inner disk with a very high scale height. In this limit the temperature is simply a sinusoidal function of $\phi$. We call this case ``wide''.
\end{enumerate}
In order to adequately investigate these scenarios we must run a total of five simulations, described in Table \ref{tab:scenarios}. For each of our two science cases we run an
additional simulation with no planet in order to subtract out effects that are purely due to the shadow. Finally, we also run a control simulation without any shadow.
Figure \ref{fig:setup} shows the initial condition of each case described here.

\begin{table}
    \begin{tabular}{ccccc}
        \hline
        \hline
        Name & $a$ & $b$ & $h$ & $m_p$ \\
        \texttt{no\_shadow} & - & - & 0 & $\mpl$ \\
        \texttt{narrow\_with} & 0.15 & 1.32 & 0.9 & $\mpl$ \\
        \texttt{narrow\_without} & 0.15 & 1.32 & 0.9 & 0 \\
        \texttt{wide\_with} & 0 & 0 & 0.9 & $\mpl$ \\
        \texttt{wide\_with} & 0 & 0 & 0.9 & 0 \\
        \hline
    \end{tabular}
    \caption{Simulation parameters. The planet mass of $\mpl$ is analogous to a planet three times as massive as Earth.}
    \label{tab:scenarios}
\end{table}

\subsection{Simulation run}
\label{subsec:sim_run}

We ran each of the five simulations for 100 orbits ($t = 200\pi$) and output the simulation results as an HDF5 file every
quarter orbit ($\Delta t = \frac{\pi}{2}$). The end state of each simulation is shown in Figure \ref{fig:end}.

\begin{figure*}
    \includegraphics[width=1.0\textwidth]{figures/final_density.pdf}
    \caption{Final gas density anomaly maps for our five simulations. Note the patterns are dominated by
    the rings created by the shadow. The radial axis is scaled logarithmically.
    }
    \label{fig:end}
    \script{plot_final.py}
\end{figure*}

\subsection{Isolating planetary effects}
It is clear in Figure \ref{fig:end} that the major deviations in our shadowed disk are due to the shadow rather than the planet. Programically determining
the location of the planet's spiral peak is a non-trivial task. First, we determine the density anomaly $\Delta \rho = \rho - \rho_0$. In the regime where all perturbations
are small we expect that
\begin{equation}
    \Delta \rho_{\rm planet} = \Delta \rho - \Delta \rho_{\rm shadow}
    \label{eq:rho_planet}
\end{equation}

$\Delta \rho_{\rm shadow}$ is determined from our shadow-only simulations (\texttt{narrow\_without} and \texttt{wide\_without}). See Figures \ref{fig:sp_narrow}
\& \ref{fig:sp_wide} for this residual in the narrow and wide cases, respectively.

Now that we have a density anomaly map we need to determine the location of the spiral peak (in $\phi$) as a function of $r$. Absent of shadow perturbations
(i.e., in the \texttt{no\_shadow} case or if the subtraction in Equation \ref{eq:rho_planet} was perfect) this would be as simple as
finding the value of $\phi$ that maximizes $\rho(\phi,r)$ for any given $r$. However, this is not true in the general case as there are additional
effects that are only found in simulations containing both the planet and shadow. In order to ensure we don't find some other maximum we must constrain
the domain of our search along the $\phi$ axis. One constraint we can utilize is that $\phi_{\rm peak}(r=1) = \phi_{\rm planet}$ as that is where the spiral is
launched from. Since the spiral is a continuous function $\phi(r)$ (that is, its pitch angle is defined, see Section \ref{subsec:spiral_morph}) we can assume that
$\phi(r_i) \approx 2\phi(r_{i-1}) - \phi(r_{i-2})$. Therefore, we can restrict our search for $\phi_{\rm peak}$ based on an initial guess. We restrict searches
to within 0.1 radians of this guess.

\begin{figure*}
    \script{plot_spirals_narrow.py}
    \includegraphics[width=1.0\textwidth]{figures/spirals_narrow.pdf}

    \caption{
        Spiral patterns observed after 99.25 orbits in the narrow shadow case. Shown is the density anomaly for every case.
        The third panel from the left shows the residuals after subtracting the first panel from the second. Plotted on the two
        rightmost panels are the locations identified as being at the peak of the spiral. Note that the color scale on the leftmost
        two panels has been scaled so that all panels can share the same colorbar.
    }
    \label{fig:sp_narrow}
\end{figure*}

\begin{figure*}
    \script{plot_spirals_wide.py}
    \includegraphics[width=1.0\textwidth]{figures/spirals_wide.pdf}
    \caption{ Same as Figure \ref{fig:sp_narrow} but for the wide shadow case.}
    \label{fig:sp_wide}
\end{figure*}

\section{Results}
\label{sec:results}

\subsection{Spiral Morphology}
\label{subsec:spiral_morph}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.5\textwidth]{figures/zeta_narrow.pdf}
        \script{plot_zeta_narrow.py}
        \caption{
            The pitch angle $\zeta$ for the narrow case. Far from the planet there is some discontinuous variability introduced by the shadow.
            The planet-only case fits well to the analytic approximation.
        }
        \label{fig:zeta_narrow}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.5\textwidth]{figures/zeta_wide.pdf}
        \script{plot_zeta_wide.py}
        \caption{
            Same as Figure \ref{fig:zeta_narrow} but for the wide case. Note that the variations in the inset are much more continuous than in the
            narrow case, and have a much larger amplitude. Variations also evident interior to the planet, but the features are not as long-lived
            as those outside.
        }
        \label{fig:zeta_wide}
    \end{center}
\end{figure}

The shape of a spiral can be parameterized by its initial condition $\phi(r_0)$ and by its pitch angle $\zeta$. Pitch angle
is defined as
\begin{equation}
    \cot \zeta = \frac{d\phi}{d \ln r}
    \label{eq:zeta_def}
\end{equation}

according to \citet{zhu2022}. We use a central difference scheme to take the numerical derivative of $\phi_{\rm peak}$ with respect to $\ln r$.
\citet{zhu2022} also provides an analytic solution in the limit that the wave has propagated far enough away from the planet:

\begin{equation}
    \zeta = \arctan |(\Omega_{\rm pat} - \Omega)r/c_{\rm s}|
\end{equation}

where $\Omega$ is the local Keplerian frequency and $\Omega_{\rm pat}$ is the Keplerian frequency at the
radius from which the spiral is launched. As we are only considering circular orbits we see that 
$\Omega_{\rm pat} = \Omega(r=1)$ and the pitch angle reduces to

\begin{equation}
    \cot \zeta = |(1-r^{-3/2}) r/c_{\rm s}|
    \label{eq:zeta_omega}
\end{equation}

We plot this analytic case against our numerical results in Figures \ref{fig:zeta_narrow} \& \ref{fig:zeta_wide}. While the simulation with
no shadow matches this solution very well we see that both shadowed cases display oscillations -- particularly far outside the planet's orbit.
We can see that if $r$ is held constant then $\zeta$ is purely a function of the sound speed $c_{\rm s}$. We can set
Equations \ref{eq:zeta_def} \& \ref{eq:zeta_omega} equal to each other to find

\begin{equation}
    \frac{d\phi}{d r} = \frac{1-r^{-3/2}}{c_{\rm s}}
\end{equation}

where we have removed the log from the differential and divided out the additional factor of $r$. We can separate the differential to see
\begin{equation}
    \int_{\phi_p}^{\phi_f} c_{\rm s} d \phi = \int_{1}^{r_f} (1-r^{-3/2}) d r
    \label{eq:analytic_shadow}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/spiral_shape_analytic.pdf}
    \caption{
        Analytic spiral curves for a variety of situations. Here we consider both the narrow and wide cases for multiple values of $h$.
        The $h=0.9$ case which we simulate numerically very nearly matches the shadowless case, shown in black. In order to get
        variations similar to those seen in our simulations we have to set $h$ to be much less than 1. The region near the planet
        has been cut out because the analytic treatment is not valid there.
    }
    \label{fig:analytic_spiral}
    \script{plot_sound_speed_analytic.py}
\end{figure}

Where $\phi_p$ is the azimuthal angle of the planet.
$c_{\rm s}(\phi)$ is non-trivial to integrate by hand, but can be done very easily numerically. We solve Equation \ref{eq:analytic_shadow}
using the Newton-Raphson method to find $\phi_f$ for a given $r_f$. Figure \ref{fig:analytic_spiral} shows that the sound speed variations
present in our simulations are insufficient to produce the spiral shapes we observe. Therefore, there must be some secondary effect not accounted for
in the analytic approximation that is causing these variations. Likely, it is some interaction between the planet's spiral and effects caused by the
shadow; we explore this in Section \ref{subsec:effect_shadow}.

\subsection{Effects from the shadows}
\label{subsec:effect_shadow}

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/slice.pdf}
    \caption{
        Density anomaly along $\phi = 0$ for every simulation output after the 25th orbit of the planet. A semi-transparent line
        is plotted for each dataset, showing that the location of the rings is static, while their amplitudes do change.
    }
    \label{fig:slice}
    \script{plot_slice.py}
\end{figure}

It can be clearly seen in Figures \ref{fig:end}, \ref{fig:sp_narrow} \& \ref{fig:sp_wide} that the morphology of the disk is dominated
by effects from the shadow, even for the very weak shadows we simulate in this work. In all simulations involving a shadow we observe a very clear
ring pattern. These rings appear after $t\sim10~P_{\rm orb}$ and are static, as shown in Figure \ref{fig:slice} It is interesting to note that
the innermost ring in the narrow case is a minimum, while in the wide case it is a maximum. Inside the innermost ring is a spiral region where gas
is interacting with the boundary.

\pagebreak

\section{Conclusion}
\label{sec:conclusion}

We have demonstrated here that even a weak shadow can induce significant structural changes in a propoplanetary disk, as well as affect the spiral path of
a planet embedded in the disk. We saw that the pitch angle of the planet's spiral was modulated by the shadow and that this modulation
was much greater than expected by an analytical treatment. To observe the refraction expected analytically we would have to reduce $h$ to be much less than 1,
and in turn we would have to increase $m_p$ greatly on order to still detect a spiral. A massive enough planet could open a gap in the disk that would interact with
the rings created by the shadow.

We also leave much of the shadow parameter space unexplored. Future investigations could study a variety of parameters, including
\begin{enumerate}
    \item The properties of the gas. We assume instant heating and cooling, but a longer timescale would mean that $c_{\rm s}$ is not
        purely a function of $\phi$ as gas in the inner portions of the disk would travel a significant fraction of its orbit before cooling completely.
    \item The radial temperature profile. We assume that $c_{\rm s}$ is not a function of $r$. However, the outer portions of the disk would realistically
        be cooler.
    \item Shadow shape. We assume shadows cast by a polar-inclined interior disk. However, if the misalignment is small then shadows on a disk can have a very different
        shape -- one that is a function of both $\phi$ and $r$.
    \item Motion of the shadow. In this study we simulated a static shadow -- analogous to that produced by an inclined interior disk. However,
        if instead a shadow is caused by an eclipsing interior binary then it would sweep around the disk on the timescale of the binary orbit.
    \item Limiting regimes of $h$. At what value do we start to see rings formed? What happens in the case of a very strong shadow?
    \item Different width regimes. At what point does the interior ring switch from a minimum to a maximum?
\end{enumerate}

\section{Acknowledgements}
This work makes use of the Athena++ MHD code version 21 \citep{athena++developmentteam2021}.
This manuscript was typeset using the open-science software \showyourwork, which makes the entire manuscript, including each figure,
completely reproducible by anyone with access to the git repository at \url{https://github.com/tedjohnson12/shadow-spirals}.

\bibliography{syw,spirals}

\end{document}